### 2020-10-02 tikz_images/proof.tex.preview.tex

%& C:\Users\tn5093\AppData\Roaming\TikzEdt\TikzEdt\023~1.0\TEMP_H~1
\begin{document}

\begin{tikzpicture}[font=\footnotesize]
% Bounding box
\draw [gray] (0,-0.25) rectangle (8,7.5);

% Outer adversary rectangle
\draw[rounded corners=15pt, fill=gray!30] (0.5,7) rectangle (7.5,2.75);
% Inner adversary rectangle
\draw[rounded corners=15pt, fill=gray!60] (1,6.5) rectangle (7,5.5);

% Adversaries and challenger
\node at (4,3.5) {\LARGE$\mathcal{A}'$};
\node at (4,6) {\LARGE$\mathcal{A}$};
\node at (4,0.5) {\LARGE$\mathcal{C}$};

% Arrows
\node[text width=3cm, align=center] at (2,1.25) {Joye-Libert\\AO scheme};
\draw [-latex] (2,1.75) -- (2,2.75);

\node[text width=3cm, align=center] at (2,4) {Our\\LCAO scheme};
\draw [-latex] (2,4.5) --(2,5.5);

\node[text width=3cm, align=center] at (6,4) {``Break''\\Our scheme};
\draw [-latex] (6,5.5) -- (6,4.5);

\node[text width=3cm, align=center] at (6,1) {``Break''\\Joye-Libert\\scheme};
\draw [-latex] (6,2.75) -- (6,1.75);

\usetikzlibrary{calc}
\pgftransformreset
\node[inner sep=0pt,outer sep=0pt,minimum size=0pt,line width=0pt,text width=0pt,text height=0pt] at (current bounding box) {};
%add border to avoid cropping by pdflibnet
\foreach \border in {0.1}
  \useasboundingbox (current bounding box.south west)+(-\border,-\border) rectangle (current bounding box.north east)+(\border,\border);
\newwrite\metadatafile
\immediate\openout\metadatafile=\jobname_BB.txt
\path
  let
    \p1=(current bounding box.south west),
    \p2=(current bounding box.north east)
  in
  node[inner sep=0pt,outer sep=0pt,minimum size=0pt,line width=0pt,text width=0pt,text height=0pt,draw=white] at (current bounding box) {
\immediate\write\metadatafile{\p1,\p2}
};
\immediate\closeout\metadatafile
\end{tikzpicture}

\end{document}

gbox (current bounding box.south west)+(-\border,-\border) rectangle (current bounding box.north east)+(\border,\border);
\newwrite\metadatafile
\immediate\openout\metadatafile=\jobname_BB.txt
\path
  let
    \p1=(current bounding box.south west),
    \p2=(current bounding box.north east)
  in
  node[inner sep=0pt,outer sep=0pt,minimum size=0pt,line width=0pt,text width=0pt,text height=0pt,draw=white] at (current bounding box) {
\immediate\write\metadatafile{\p1,\p2}
};
\immediate\closeout\metadatafile
\end{tikzpicture}

\end{document}
kzpicture}

\end{document}

ocument}

ment}
t}