### 2022-02-01 reviews/TAC_feb_2022/COMBINED REVIEWS AND EDITOR RESPONSE.txt
TAC conditional accept received. Added reviewer comments Added new author's response document (possibly to be renamed to document of changes) Added files for creating latexdiff between accepted and final submission

COMBINED REVIEWS AND EDITOR RESPONSE

--------------------------------------------------------------------------------
EDITOR
--------------------------------------------------------------------------------

Dear Dr. Ristic:

Your paper “Distributed Range-Only Localisation that Preserves Sensor and Navigator Privacies”, Number 21-1548, that you submitted for possible publication as a Regular Paper in the IEEE Transactions on Automatic Control has now been reviewed.  Copies of the reviews are enclosed.

I am pleased to inform you that the paper is acceptable for publication in the Transactions provided that you can make the modifications described below.

The reviewers would like the authors to clarify some concerns such as the assumption of the EKF design, private sensor variance information, and research motivation and challenge, and so forth.

As soon as you have completed these modifications, please submit your revised paper. Please include a detailed statement describing the changes in the manuscript. The revised paper must be received within 5 weeks of the date of this letter.

Please log-in, click on the "Author" link at the bottom of your access page, and upload your revised paper under the SAME reference number.

I look forward to receiving your revised paper.

Sincerely,

Dr. Zhiwei Gao
Associate Editor
zhiwei.gao@northumbria.ac.uk, zhiwegac@gmail.com Northumbria University

--------------------------------------------------------------------------------
REVIEWER 1
--------------------------------------------------------------------------------

Thank you for updating the manuscript and most of the concerns have
been addressed well in the response letter. However, there is one
problem still confused me. The authors claimed that the Gaussian
assumption has been removed in the revised version, however, the EKF
(EIF) was still used for prediction (Eqs. 26-27). Basically, the EKF is
based on the Gaussian assumption and the prediction is obtained in the
sense of mean value. Then, the non-Gaussian dynamics would affect the
prediction performance using EKF if the Gaussian assumption is removed.
My point is 1) if the assumption is removed, please explain the
non-Gaussian influence from the EKF or 2) if the Gaussian assumption
remains in the manuscript, please explain from the noise comes from in
physics sense.

--------------------------------------------------------------------------------
REVIEWER 2
--------------------------------------------------------------------------------

The authors have significantly improved the readability of the paper.
The contribution is significant enough for privacy preserving state
estimation. However, this reviewer still have few comments:
1- It could be interesting to comment why sensor variance is a private
information. Why is it crucial to preserve it if measurements are
already protected?

2- In the sentence 'homomorphic encryption is used to make
time-independent model-free location estimates where an estimator does
not learn sensor measurements or locations.; Shouldn't be 'cannot
learn'?

3- In problem statement 'we consider the context of privacy-preserving
range sensor navigation, where we want no sensor to learn information
about the navigator or other sensors beyond their local measurements,
and the navigator to learn no information about individual sensors
beyond its location estimate.', please specify type of information.

4- Indistinguishable weights: what would happen if the sensor learns
navigator weights?

5- Fix the following sentence: 'If an attacker compromises the
navigator, they have control over the weights,'

6- Define IND-CPA.

7- Simulation: The authors have found that 1.7 s are needed for each
computation step, is that compatible with real-time operation?

8- Can you comment why quantization noise, involved in the
encryption-decryption schemes, isn't taken into account in the filter
equations?

--------------------------------------------------------------------------------
REVIEWER 3
--------------------------------------------------------------------------------

Comments:
This paper proposes a novel distributed localisation method in the
presence of range-only sensors, which preserves both navigator and
sensor privacies. The major contribution of this paper is that a novel
private linear combination aggregation scheme is proposed, and based on
that, a modified extended Kalman filter is also derived. Some comments
are given as follows:

1.In this paper, the full names of some abbreviations are not given.
For example, in Section I, page 1, left column, “AES”and “RSA”, and in
Section I, page 1, right column, “pWSAc” and “pWSAh”.

2.The authors should elaborate the advantages and disadvantages of the
existing typical cryptographic secrecy scheme and the motivation for
proposing private linear combination aggregation scheme in this paper.

3.Some symbols are reused. For example, In Section I, Notation,
“timestep k” and “will denote encryption and decryption with key k”.

4.The authors need to present theoretical computational complexity of
the proposed method.

5.The authors should state how to measure the performance of the
private linear combination aggregation scheme proposed in this paper.

6.In practical engineering application, how to balance the relationship
between key sizes and computation.